---
title: "Geohashing in R"
author: "Oliver Keyes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geohashing in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*geohashes* are a way of representing latitude/longitude pairs as individual, short strings of numbers and letters. The <code>geohash</code> package provides tools for creating and manipulating these.

## Encoding and decoding

To turn a set of latitude/longitude pairs into a geohash, you need the <code>gh_encode</code> function. This accepts latitude, longitude and precision arguments - where precision refers to the number of characters the hash should take up and (correspondingly) how accurate it is. Any precision between 1 and 10 is acceptable - something outside that range will produce a warning, and result in a default of 6 being used.

```{r, eval=FALSE}
# Encode!
gh_encode(42.60498046875, -5.60302734375, 5)
# [1] "ezs42"
```

To decode, you use <code>gh_decode</code>, which works the other way around - it accepts hashes, and returns a data.frame of latitude, longitude, and the error rates for each (since some degree of inaccuracy is introduced by the hashing process):

```{r, eval=FALSE}
# Decode!
gh_decode("ezs42")
# lat      lng      lat_error  lng_error
# 42.60498 42.60498 0.02197266 0.02197266
```

Both functions are fully vectorised, so if you've got an entire set of latitudes and longitudes (or an entire set of hashes) it's not a problem! They're based on compiled code and can handle a million lat/long pairs (or hashes) in a third of a second.

## Neighbours

Because geohashes are calculated for geographic "boxes", they have neighbours; a hash that represents the block to the north, or south, or northwest, or...so on and so forth. You can extract these in one of two ways. With <code>gh_neighbours</code>, it'll extract every hash around the hash you provide:

```{r, eval=FALSE}
# Get all neighbours!
gh_neighbours("ezs42")
#  north northeast  east southeast south southwest  west northwest
#  ezs48     ezs49 ezs43     ezs41 ezs40     ezefp ezefr     ezefx
```

If you only want a specific neighbour, no problem! You can use convenience functions named after the cardinal and intercardinal points (<code>north()</code>, <code>northwest()</code> etc) to grab individual neighbours:

```{r, eval=FALSE}
# Get the southeast neighbours!
southeast("ezs42")
# [1] "ezs41"
```

These functions are slower than encoding and decoding, since they're a lot more complex, but still pretty fast.

## Feature ideas or issues
If you have ideas for other geohash-related features, or run into a bug, the best approachis to either [request/report it](https://github.com/Ironholds/geohash/issues) or [add it](https://github.com/Ironholds/geohash/pulls)!
